PM Mungbean meta-analysis

### 2020-05-22 01_Meta-analysis_experiments.Rmd

**2001 - 2004 trials**  
Trial data were obtained from a non-published manuscript.  
Ratings for powdery mildew were on a 1-5 rating scale and can be read in the file "Powdery mildew paper#1.doc".
Comparing trials, which have not used the same rating scale will be done using a percentage change.

>>>>

**2001 - 2004 trials**  
Trial data were obtained from a non-published manuscript.  
Ratings for powdery mildew were on a 1-5 rating scale where 1 = no sign of disease, 2 = 1-30% diseased leaf area, 3 = 31-65% diseased leaf area, 4 = 66-99% diseased leaf area and 5 = complete defoliation.  
Comparing trials which have not used the same rating scale will be done using a
percentage change.

### 2020-05-22 02_Preliminary_analysis.Rmd

Various factors have been studied in the previous experiments which may influence the following results. However unless stated the following plots and analyses were made disregarding effects from other factors not mentioned.

   1) Fungicide type

   2) Fungicide dose (varies only slightly within some fungicide types)

   3) Number of fungicide sprays

   4) Timing of fungicide spray/s relative to first sign of disease (_i.e_. when pathogen is present)

   5) Host cultivars (probably is a co-variate with season due to changing cultivars over time)

   6) Experiment location

   7) Row spacing

   8) In-crop rainfall

   9) Irrigated vs non-irrigated trials

   10) Mean daily temperature in the weeks following spray application

   11) Final disease rating

   12) Two pathogen species cause PM (new discovery in 2018)

The data plotted below are from `r length(unique(PM_MB_means$trial_ref))` field trials between (`r first(levels(PM_MB_means$year))` - `r last(levels(PM_MB_means$year))`) of which the details are described in [Studies considered in meta-analysis section](#studies-considered-in-meta-analysis).  
All data explored below are a list of mean values from each treatment applied in each experiment. Some experiments have in-season observations of disease incidence and severity, however here we have summarised the data to show only the end of season observations.

### Fungicides
#### Fungicide type
Let's explore how many different types of fungicide were used and at what frequency.

@@

PM_MB_means %>%
   mutate(fungicide_ai = as.factor(PM_MB_means$fungicide_ai)) %>%
   filter(total_fungicide == 1 | total_fungicide == 2) %>%
   filter(
      fungicide_ai != "pyrazophos" &
         fungicide_ai != "control" &
         fungicide_ai != "benomyl" &
         fungicide_ai != "Acibenzolar-S-methyl"
   ) %>%
   mutate(fungicide_ai = factor(
      fungicide_ai,
      levels = c(
         "carbendazim",
         "pyraclostrobin",
         "sulphur",
         "200 g/L azoxystrobin +\n80 g/L cyproconazole",
         "tebuconazole",
         "propiconazole",
         "200 g/L tebuconazole +\n120 g/L azoxystrobin"
      )
   )) %>%
   ggplot(aes(y = prop_YG, x = fungicide_ai)) +
   facet_grid(rows = vars(total_fungicide)) +
   geom_boxplot(aes(fill = fungicide_ai, alpha = 0.5)) +
   geom_hline(aes(yintercept = 0), size = 0.5) +
   labs(y = "Proportion of grain yield saved",
        title = "Grain yield proportion saved\ngrouped by fungicide type") +
   theme(legend.position = "none") +
   scale_fill_usq() +
   coord_flip()
```

@@

```{r PropiDoseEffect}
PM_MB_means %>%
   filter(fungicide_ai == "propiconazole") %>%
   ggplot(aes(x = relevel(as.factor(dose_ai.ha), "62.5"), y = prop_YG)) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Yield for each respective propiconazole dose") +
   geom_boxplot() +
   scale_fill_usq() +
   scale_colour_usq()
```

@@

This plot plot, however, indicated the previous plot might be skewed by high yielding trials.

@@

```{r Host_genotype}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(host_genotype, fungicide_ai, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Treatments = n) %>%
   ggplot(aes(x = host_genotype, y = Treatments, fill = fungicide_ai)) +
   xlab("Cultivar") +
   ylab("N Trials") +
   ggtitle(label = "Cultivars used in either tebuconazole or propiconazole trials") +
   geom_col() +
   scale_fill_usq()
```

@@

First let's look at overall yields for each cultivar, where trial is not a factor, for all treatments (spray and no spray).  
```{r yield_vol}
# this could be converted into a shiny app as a chunk. Make each input for the function a shiny option
# Outputs the table and graph
source("R/yield_volatility.r") #function to investigate the volatility in yields
# Volitility is the range between the upper and lower 97.5 and 2.5 % quartiles

@@

YV1[[3]]
```
Now let's look at overall yields for each cultivar, where trial is not a factor, in only the no spray treatments.
```{r yieldVol_control}

@@

Let's examine the same plots as above but only for the most commonly used trial site, Hermitage.  
```{r yieldVol_Herm}
YV1_herm <-
   yield_volatility(
      genotype_by_trial = FALSE,
      control_only = FALSE,
      location = "Hermitage"
   )
YV1_herm[[2]] +
   scale_fill_usq()
```

@@

Let's look at Jade, to see if it differs.
```{r YieldVol_Jade}
YV2_crystal <- yield_volatility("Jade",control_only = FALSE)
YV2_crystal[[2]] +
   scale_fill_usq()
```

For Jade there is high variation between trials and with in. Perhaps indicating lower tolerance. Let's look at the same plot with data only from no spray controls to see if the variation is the same.

@@

```{r RS_severity}
# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot() +
   ggtitle("Powdery mildew variation between different row spacing")
```

```{r RS_severity_x_trial}
# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot() +
   facet_grid(cols = vars(location)) +
   ggtitle("Powdery mildew variation between different row spacing")
```

@@

```{r RS_yield}
# Which row spacing leads to the higher yield potential
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   ggplot(aes(
      y = as.numeric(grain_yield.t.ha.),
      x = row_spacing,
      colour = location
   )) +
   geom_jitter(width = 0.01) +
   ggtitle("Grain yield results for different row spacing at three locations") +
   xlab("Row spacing (m)") +
   ylab("Grain yield - tonnes per hectare") +
   scale_colour_usq()
```

@@

To visualise the model fits we will use a heat map to describe how the models faired in relation to the rainfall windows. log of the p value was used to improve the resolution of small very small p values.  
```{r rainLMpvalue}
ggplot(lm_rain, aes(x = start_day, y = end_day, z = log(lm_pval))) +
   geom_raster(aes(fill = log(lm_pval)), interpolate = TRUE) +
   geom_contour(bins = 15, colour = "white") +
   geom_point(data = lm_rain[which(lm_rain$lm_pval == min(lm_rain$lm_pval)), ],
              aes(x = start_day, y = end_day),
              colour = " red") +
   scale_fill_distiller(palette = "RdBu", direction = -1) 
```

@@

**Fungicide effect on proportional yield gain with respect to time of application**  
First let's look at the quantiles for each fungicide irrespective of other factors.
We will focus on proportional yield gain, compared to no treatment control, as this should partially correct for some spatial and seasonal effects.  
```{r fungicide_proportion_yield}
PM_MB_means %>%
   group_by(fungicide_ai) %>%
   summarise(
      n = length(prop_YG),
      lower_2.5 = quantile(prop_YG, na.rm = T, c(0.025)),
      median = median(prop_YG, na.rm = T),
      mean = mean(prop_YG, na.rm = T),
      upper_97.5 = quantile(prop_YG, na.rm = T, c(0.975))
   ) %>%
   filter(n >= 5) %>% # remove any fungicide groups with less than 5 observations
   arrange(desc(median)) %>%
   kable(caption = "Fungicide effect on proportion of yield gain") %>%
   kable_styling(fixed_thead = T, full_width = T)

# plot table as a box plot
PM_MB_means %>%
   filter(
      fungicide_ai == c(
         "200 g/L tebuconazole +\n120 g/L azoxystrobin" ,
         "200 g/L azoxystrobin +\n80 g/L cyproconazole",
         "tebuconazole",
         "propiconazole",
         "sulphur"
      )
   ) %>%
   ggplot(aes(y = prop_YG, x = fungicide_ai)) +
   geom_boxplot() +
   geom_hline(aes(yintercept = 0), size = 0.5, colour = "red") +
   labs(x = "Proportion of grain yield saved",
        title = "Grain yield proportion saved grouped by\nfungicide type") +
   theme(
      panel.grid.major = element_line(colour = "grey80"),
      panel.grid.minor = element_line(colour = "grey50"),
      panel.background = element_rect(fill = NA)
   ) +
   guides(fill = guide_legend(title = "Fungicide")) +
   coord_flip()
```

>>>>

Various factors have been studied in the previous experiments which may influence the following results.
However, unless stated, the following plots and analyses were made disregarding effects from other factors not mentioned.

   1 Fungicide type

   2 Fungicide dose (varies only slightly within some fungicide types)

   3 Number of fungicide sprays

   4 Timing of fungicide spray/s relative to first sign of disease

   5 Host cultivars (probably is a co-variate with season due to changing cultivars over time)

   6 Experiment location

   7 Row spacing

   8 In-crop rainfall

   9 Irrigated vs non-irrigated trials

   10 Mean daily temperature in the weeks following spray application

   11 Final disease rating

   12 Two pathogen species cause PM (new discovery in 2018)

The data plotted below are from `r length(unique(PM_MB_means$trial_ref))` field trials between (`r first(levels(PM_MB_means$year))` - `r last(levels(PM_MB_means$year))`) of which the details are described in [Studies considered in meta-analysis section](#studies-considered-in-meta-analysis).

All data explored below are a list of mean values from each treatment applied in each experiment.
Some experiments have in-season observations of disease incidence and severity.
However, here we have summarised the data to show only the end of season observations.

### Fungicides

#### Fungicide type

Let's explore how many different types of fungicide were used and at what frequency.

@@

PM_MB_means %>%
   mutate(fungicide_ai = as.factor(PM_MB_means$fungicide_ai)) %>%
   filter(total_fungicide == 1 | total_fungicide == 2) %>%
   filter(
      fungicide_ai != "pyrazophos" &
         fungicide_ai != "control" &
         fungicide_ai != "benomyl" &
         fungicide_ai != "Acibenzolar-S-methyl"
   ) %>%
   mutate(fungicide_ai = factor(
      fungicide_ai,
      levels = c(
         "carbendazim",
         "pyraclostrobin",
         "sulphur",
         "200 g/L azoxystrobin +\n80 g/L cyproconazole",
         "tebuconazole",
         "propiconazole",
         "200 g/L tebuconazole +\n120 g/L azoxystrobin"
      )
   )) %>%
   ggplot(aes(y = prop_YG, x = fungicide_ai)) +
   facet_grid(rows = vars(total_fungicide)) +
   geom_boxplot(aes(fill = fungicide_ai)) +
   geom_hline(aes(yintercept = 0), size = 0.5) +
   labs(y = "Proportion of grain yield saved",
        title = "Grain yield proportion saved\ngrouped by fungicide type") +
   theme(legend.position = "none") +
   scale_fill_usq() +
   coord_flip()
```

@@

```{r PropiDoseEffect}
PM_MB_means %>%
   filter(fungicide_ai == "propiconazole") %>%
   ggplot(aes(x = relevel(as.factor(dose_ai.ha), "62.5"), y = prop_YG)) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Yield for each respective propiconazole dose") +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5)
```

@@

This plot, however, indicates that the previous plot might be skewed by high yielding trials.

@@

```{r Host_genotype}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(host_genotype, fungicide_ai, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Treatments = n) %>%
   ggplot(aes(x = host_genotype, y = Treatments, fill = fungicide_ai)) +
   xlab("Cultivar") +
   ylab("N Trials") +
   ggtitle(label = "Cultivars used in either tebuconazole or propiconazole trials") +
   geom_col() +
   scale_fill_usq(name = "Fungicide AI")
```

@@

First let's look at overall yields for each cultivar, where trial is not a factor, for all treatments (spray and no spray).

```{r yield_vol}
# this could be converted into a shiny app as a chunk. Make each input for the function a shiny option
# Outputs the table and graph
source("R/yield_volatility.r") #function to investigate the volatility in yields
# Volitility is the range between the upper and lower 97.5 and 2.5 % quartiles

@@

YV1[[3]]
```

Now let's look at overall yields for each cultivar, where trial is not a factor, in only the no spray treatments.

```{r yieldVol_control}

@@

Let's examine the same plots as above but only for the most commonly used trial site, Hermitage.

```{r yieldVol_Herm}
YV1_herm <-
   yield_volatility(
      genotype_by_trial = FALSE,
      control_only = FALSE,
      location = "Hermitage"
   )
YV1_herm[[2]] +
   scale_fill_usq()
```

@@

Let's look at Jade, to see if it differs.

```{r YieldVol_Jade}
YV2_crystal <- yield_volatility("Jade",control_only = FALSE)
YV2_crystal[[2]] +
   scale_fill_usq()
```

For Jade there is high variation between trials and with-in.
Perhaps indicating lower tolerance. Let's look at the same plot with data only from no spray controls to see if the variation is the same.

@@

```{r RS_severity}
# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   ggtitle("Powdery mildew variation between different row spacing")
```

```{r RS_severity_x_trial}
# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   facet_grid(cols = vars(location)) +
   ggtitle("Powdery mildew variation between different row spacing") +
   ylab("Final severity rating") +
   xlab("Row spacing (m)")
```

@@

```{r RS_yield}
# Which row spacing leads to the higher yield potential
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   ggplot(aes(
      y = as.numeric(grain_yield.t.ha.),
      x = row_spacing,
      colour = location
   )) +
   geom_jitter(width = 0.01) +
   ggtitle("Grain yield results for different row spacing at three locations") +
   xlab("Row spacing (m)") +
   ylab("Grain yield (t/ha)") +
   scale_colour_usq()
```

@@

To visualise the model fits we will use a heat map to describe how the models faired in relation to the rainfall windows.
log of the p value was used to improve the resolution of small very small p values.

```{r rainLMpvalue}
ggplot(lm_rain, aes(x = start_day, y = end_day, z = log(lm_pval))) +
   geom_raster(aes(fill = log(lm_pval)), interpolate = TRUE) +
   geom_contour(bins = 15, colour = "white") +
   geom_point(data = lm_rain[which(lm_rain$lm_pval == min(lm_rain$lm_pval)), ],
              aes(x = start_day, y = end_day),
              colour = " red") +
   scale_fill_distiller(palette = "RdBu", direction = -1) 
```

@@

**Fungicide effect on proportional yield gain with respect to time of application**  
First let's look at the quantiles for each fungicide irrespective of other factors.
We will focus on proportional yield gain, compared to no treatment control, as this should partially correct for some spatial and seasonal effects.

```{r fungicide_proportion_yield}
# plot table as a box plot
PM_MB_means %>%
   filter(
      fungicide_ai == c(
         "200 g/L tebuconazole +\n120 g/L azoxystrobin" ,
         "200 g/L azoxystrobin +\n80 g/L cyproconazole",
         "tebuconazole",
         "propiconazole",
         "sulphur"
      )
   ) %>%
   ggplot(aes(y = prop_YG, x = fungicide_ai)) +
   geom_boxplot(aes(fill = fungicide_ai)) +
   geom_hline(aes(yintercept = 0), size = 0.5, colour = "red") +
   labs(x = "Proportion of grain yield saved",
        title = "Grain yield proportion saved,\ngrouped by fungicide type") +
   theme(
      panel.grid.major = element_line(colour = "grey80"),
      panel.grid.minor = element_line(colour = "grey50"),
      panel.background = element_rect(fill = NA)
   ) +
   guides(fill = guide_legend(title = "Fungicide")) +
   scale_fill_usq() +
   coord_flip()
```
